Active Code Completion

### 2012-03-19 icse12/graphite-icse12.tex
Shortened some of the sentences.

\begin{figure*}\label{color}
\begin{center}
\includegraphics[scale=0.8]{color_palette.png}\end{center}
\caption{(a) An example code completion palette associated with the \texttt{Color} class. (b) The source code generated by this palette.}
\end{figure*}

@@

The developer can interact with such palettes to provide parameters and other information related to her intent, and receive immediate feedback about the effect these choices will have on the behavior of the object being constructed. When this interaction is complete, the palette is responsible for generating appropriate source code for insertion at the cursor. Figure 1b shows the code that is inserted after the user presses the \textsc{enter} key.

@@

Next, we describe Graphite, an Eclipse plug-in that provides active code completion for users of the Java programming language (Section 5). We describe several interesting design choices that were made to satisfy many of the requirements discovered in our preliminary investigations and briefly examine the trade-offs that were necessary due to Java's purely textual program representations and Eclipse's limited extension APIs.

@@

In addition to asking for a simple rating for each palette, we also solicited open-ended comments. A large number of participants volunteered comments: 193 for the color palette, 129 for the regular expression palette and 142 for the SQL palette. These responses were highly valuable when developing the design criteria below and helped to explain the patterns observed in Fig. \ref{frequencies}.

@@

\subsection{Maintaining Separation of Concerns (183)}
The most common issue developers had, particularly with the Color palette, was that it seemed to ignore the issue of separation of concerns by allowing developers to insert color constants directly into the program logic. Many developers noted that this is considered bad practice, or should be limited to the prototyping phase of a project. This concern was also expressed in responses to the SQL palette, which required inserting parameters to connect to a database so that the query could be tested. The code that was inserted included initialization steps needed to connect to the specific database that was entered, and several participants noted that much of this information should appear in an external resource file rather than inline with the program logic itself. Few participants made similar comments about the regular expression palette, however, indicating that regular expressions are considered a part of the program logic rather than data by most developers.

@@

\subsection{Palette API}
In order to allow palettes to communicate with the surrounding editor environment, a simple Javascript interface was injected. These functions were implemented by the editor plug-in, but appear as ``native'' browser functions to Javascript code. The API consists of the following methods, accessed through a global object named \verb|graphite|:
\begin{itemize}
	\item \verb|insert(str)|

	Inserts the specified string at the cursor and closes the palette. The appropriate amount of indentation is automatically inserted after any newlines in this string.

	\item \verb|cancel()|

	Closes the palette without inserting a string.

	\item \verb|getSelectedText()|

	Returns the text that is currently selected by the user in the editor, or an empty string if no text has been selected. This method is used to implement reinvocation.

	\item \verb|getIDE()|

	Returns a string that specifies the IDE that is being used, ``Eclipse'' in our implementation.

	\item \verb|getLanguage()|

	Returns a string that specifies the programming language that is being used, ``java'' in our implementation.
\end{itemize}

@@

Three of the four control subjects experienced significant difficulties associated with escape sequences. One subject recognized the error fairly quickly after Eclipse complained that the escape sequences used in the pattern were invalid (they were, in fact, valid escape sequences for regular expression patterns, but not for string literals.) The subject continued to miss escape sequences occasionally throughout the remainder of the study, but was able to fix them quickly after Eclipse alerted the subject to the error. For the other two subjects, the problem was more severe. In both cases, they noticed the error that Eclipse gave, but thought it indicated a problem with their pattern itself. One subject thought that the uses of symbolic escape sequences like \verb|\(| were incorrect. He decided to replace these with ASCII escape sequences (\verb|\050|) after looking up an ASCII conversion table. This was an unnecessary and overly complex solution to the problem. The other subject thought that the problem was in his use of the whitespace escape sequence, \verb|\s|. Thus, he  spent several minutes looking up how to match whitespace correctly. In doing so, he found a description of the double-escape problem in Java and was able to fix the problem after that.
\subsection{Treatment}
One subject was already familiar with an external tool, regexpal.com (also used by a subject in the control group), and used it variously throughout the task. In most cases, the subject used the external tool's quick reference documentation while using our palette for authoring and testing. At other times, the subject used the external tool itself. The subject indicated that in some cases, he was not sure that our palette was free of bugs (when something unexpectedly matched or did not match\footnote{We did not observe any actual errors related to regex matching in our tool.}), and also because the external tool provided syntax and substring highlighting, a useful feature for complex regular expressions. When done with the external tool, the subject pasted the pattern into our tool to generate the appropriate code. As such, the subject had no difficulties with escaping or factory pattern instantiation.

The other two subjects in the treatment group did not use any external tools. Both subjects decided not to use our palette for the first task (likely because they forgot that it was available, since they spent some time looking up API documentation after our initial demo with the color palette.) They both recognized the need for factory pattern instantiation and also both had an initial error related to escaping that they were able to resolve before moving on. One of these subjects also began to write tests within a \verb|main| method.

>>>>

\begin{figure*}\label{color}
\begin{center}
\includegraphics[width=\textwidth]{color_palette.png}\end{center}
\caption{(a) An example code completion palette associated with the \texttt{Color} class. (b) The source code generated by this palette.}
\end{figure*}

@@

The developer can interact with such palettes to provide parameters and other information related to her intent, and receive immediate feedback about the effect these choices will have on the behavior of the object being constructed. When this interaction is complete, the palette generates appropriate source code for insertion at the cursor. Figure 1b shows the inserted code after the user presses the \textsc{enter}.

@@

Next, we describe Graphite, an Eclipse plug-in that provides active code completion for the Java programming language (Section 5). We describe several design choices that were made to satisfy many of the requirements discovered in our preliminary investigations and briefly examine the trade-offs that were necessary due to Java's purely textual program representations and Eclipse's limited extension APIs.

@@

In addition to asking for a simple rating for each palette, we also solicited open-ended comments. A large number of participants volunteered comments: 193 for the color palette, 129 for the regular expression palette and 142 for the SQL palette. These responses were highly valuable when developing the design criteria below and helped to explain the patterns observed in Figure \ref{frequencies}.

@@

\subsection{Maintaining Separation of Concerns (183)}
The most common issue developers had, particularly with the Color palette, was that it ignored the issue of separation of concerns by allowing developers to insert color constants directly into the program logic. Many developers noted that this is considered bad practice, or should be limited to the prototyping phase of a project. This concern was also expressed in responses to the SQL palette, which required inserting parameters to connect to a database so that the query could be tested. The resulting code included initialization steps needed to connect to the specific database that was entered, and several participants noted that much of this information should appear in an external resource file separated from the program logic. Few participants made similar comments about the regular expression palette, however, indicating that regular expressions are considered a part of the program logic rather than data by most developers.

@@

\subsection{Palette API}
In order to allow palettes to communicate with the surrounding editor environment, a simple Javascript interface was injected. These functions were implemented by the editor plug-in, but appear as ``native'' browser functions to Javascript code. The API consists of the following methods, accessed through a global object named \verb|graphite|:
\begin{itemize}
	\item \verb|insert(str)|: Inserts the specified string at the cursor and closes the palette. The indentation is automatically inserted after any newlines in this string.

	\item \verb|cancel()|: Closes the palette without changes.

	\item \verb|getSelectedText()|: Returns the text that is currently selected by the user in the editor, or an empty string if no text has been selected. This method is used to implement reinvocation.

	\item \verb|getIDE()|: Returns a string that specifies the IDE that is being used, ``Eclipse'' in our implementation.

	\item \verb|getLanguage()|: Returns a string that specifies the programming language that is being used, ``java'' in our implementation.
\end{itemize}

@@

Three of the four control subjects experienced significant difficulties associated with escape sequences. One subject recognized the error fairly quickly after Eclipse complained that the escape sequences used in the pattern were invalid (they were, in fact, valid escape sequences for regular expression patterns, but not for string literals.) The subject continued to miss escape sequences occasionally throughout the remainder of the study, but was able to fix them quickly after Eclipse alerted the subject to the error. For the other two subjects, the problem was more severe. In both cases, they noticed the error that Eclipse gave, but thought it indicated a problem with their pattern itself. One subject thought that the uses of symbolic escape sequences like \verb|\(| were incorrect. He decided to replace these with ASCII escape sequences (\verb|\050|) after looking up an ASCII conversion table. This was an unnecessary and overly complex solution to the problem. The other subject thought that the problem was in his use of the whitespace escape sequence, \verb|\s|. Thus, he  spent several minutes looking up how to match whitespace correctly. In doing so, he found a description of the double-escape problem in Java and could fix the problem after that.
\subsection{Treatment}
One subject was already familiar with an external tool, regexpal.com (also used by a subject in the control group), and used it variously throughout the task. In most cases, the subject used the external tool's quick reference documentation while using our palette for authoring and testing. At other times, the subject used the external tool itself. The subject indicated that in some cases, he was not sure that our palette was free of bugs (when something unexpectedly matched or did not match\footnote{We observed no actual errors related to regex matching in our tool.}), and also because the external tool provided syntax and substring highlighting, a useful feature for complex regular expressions. When done with the external tool, the subject pasted the pattern into our tool to generate the appropriate code. As such, the subject had no difficulties with escaping or factory pattern instantiation.

The other two subjects in the treatment group did not use any external tools. Both subjects decided not to use our palette for the first task (likely because they forgot that it was available, since they spent some time looking up API documentation after our initial demo with the color palette.) They both recognized the need for factory pattern and also both had an initial error related to escaping that they were able to resolve before moving on. One of these subjects also began to write tests within a \verb|main| method.
